Iconicity analysis first exporation
========================================================

This document presents an exploratory analysis of the iconicity data using both lists:

* List 1: the original norms by Lynn, Marcus and Gary (N=592)
* List 2: the new norms (N=1360)

Several analyses were performed:

* part-of-speech differences for List 1 and List 2
* the relationship between AOA ***ratings*** and iconicity (a conceptual replication of your submitted paper?)
* the relationship between concreteness and iconicity
* the relationship between affect and iconicity
* the relationship between sensory modality ratings and iconicity
* finally: does iconicity predict unique RT variance?

```{r echo=FALSE}
     setwd("/Users/teeniematlock/Desktop/research/iconicity/analysis/")
     options(warn=-1)
```


Load in data and packages. Relevel for better display:

```{r cache=FALSE}
     icon <- read.csv("iconicity_ratings_both.csv")
     library(ggplot2)
     library(mgcv)
     icon$ContentPOS <- factor(icon$ContentPOS,
                              levels=c("Verb","Noun","Adjective","Adverb"))
```

<br><br><br>
## POS differences

How does iconicity differ depending on parts of speech differences?

```{r cache=FALSE, fig.width=8, fig.height=4}
     ggplot(icon[!is.na(icon$ContentPOS),],
          aes(x=ContentPOS,y=Written,fill=ContentPOS)) +
          geom_boxplot() + facet_grid(~ListIdentifier)
     anova(lm(Written ~ ContentPOS,subset(icon,ListIdentifier=="list1")))
     anova(lm(Written ~ ContentPOS,subset(icon,ListIdentifier=="list2")))
```

Notice that I have omitted "grammatical" for now. This is not straightforward, e.g., how should certain adverbs be treated?

<br><br><br>
## AOA ratings

How does iconicity depend on AOA ratings? (Taken from Kuperman et al.'s mega-norming study)

```{r cache=FALSE, fig.width=8, fig.height=4}
     ggplot(icon,
            aes(x=AOA,y=Written)) +
          geom_point(shape=16) +
          geom_smooth(method="lm") + facet_grid(~ListIdentifier)
```

<br>
As expected given your search criteria, the original list has much higher AOA ratings. Overall, there is a significant downward trend with words rated to be acquired later having less iconicity, even after controlling for frequency:

```{r cache=T}
     summary(lm(Written ~ AOA+WordFreq,icon))
```

This shows the pure effect of AOA ratings with word frequency residualized out:

```{r cache=FALSE, fig.width=8, fig.height=4}
     icon$res <- NA
     icon[!is.na(icon$WordFreq),]$res <- residuals(lm(Written ~ WordFreq,icon))
     ggplot(icon,
            aes(x=AOA,y=res)) +
          geom_point(shape=16) +
          geom_smooth(method="lm")
```

<br><br><br>
## Concreteness

How is iconicity related to concreteness?

```{r cache=FALSE, fig.width=8, fig.height=4}
     ggplot(icon,
            aes(x=Conc,y=Written)) +
          geom_point(shape=16) +
          geom_smooth()
```

It seems to be that words of intermediate concreteness are the most iconic words. We can test this by fitting a quadratic model:

```{r cache=T}
     anova(lm(Written ~ Conc + I(Conc^2),icon))
```

<br><br><br>
## Affect ratings

How is iconicity related to affect? (FYI: I am taking a measure of "absolute valence" here. According to this measure highly valenced words have high numbers regardless of positive or negative valence)

```{r cache=FALSE, fig.width=8, fig.height=4}
     ggplot(icon,
            aes(x=AbsValence,y=Written)) +
          geom_point(shape=16) +
          geom_smooth()
```

No clear trend detectable and no significant effect (fitting a GAM here)

```{r cache=T}
     summary(gam(Written ~ s(AbsValence),data=icon))
```

<br><br><br>
## Iconicity and modality

For the Lynott and Connell (2009) adjective norms, let's check the relationship between iconicity and the dominant modality, e.g., "shiny" has dominant modality = "visual".

```{r cache=FALSE, fig.width=8, fig.height=4}
     ggplot(icon[!is.na(icon$DominantModality),],
            aes(x=DominantModality,y=Written,fill=DominantModality)) +
          geom_boxplot()
```

Let's do the same thing for the van Dantzig et al. (2013) norms, which are largely overlapping with Lynott and Connell (2009), but they stem from an independent rating:

```{r cache=FALSE, fig.width=8, fig.height=4}
     ggplot(icon[!is.na(icon$DantzigDominantModality),],
            aes(x=DantzigDominantModality,y=Written,
                fill=DantzigDominantModality)) +
          geom_boxplot()
```

We can test whether there are significant iconicity differences between words for the different modalities:

```{r cache=FALSE}
     anova(lm(Written ~ DominantModality,icon))
     anova(lm(Written ~ DantzigDominantModality,icon))
```


<br><br><br>
## Does iconicity predict unique variance in reaction times?

For now, I am simply testing whether iconicity predicts unique variance in reaction times above and beyond AOA and word frequency. I am going to use the Lexical Decision times from the English Lexicon Project (Balota et al., 2007). 

```{r cache=FALSE}
     summary(lm(RT ~ AOA + WordFreq + Written,icon))
```

Once AOA and WordFreq is controlled for, there seems to be no effect of iconicity. However, what about the word naming reaction times?


```{r cache=FALSE}
     summary(lm(NamingRT ~ AOA + WordFreq + Written,icon))
```

Let's visualize that effect by residualizing AOA and WordFreq out:

```{r cache=T, fig.width=8, fig.height=4}
     icon$res <- NA
     NAs <- (is.na(icon$WordFreq)|is.na(icon$AOA)|is.na(icon$NamingRT))
     icon[!NAs,]$res <- residuals(lm(NamingRT ~ WordFreq + AOA,
                                   icon[!NAs,]))
     ggplot(icon,
            aes(x=Written,y=res)) +
          geom_point(shape=16) +
          geom_smooth(method="lm") 
```
<br>
The effect is small, but hey... ~4ms less per increase of iconicity 1. Not too bad.



